---
title: "Farnham Food Hygiene Ratings"
subtitle: "Data Profiling"
author: "Clare Gibson"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_depth: 4
    toc_float: true
---

# Introduction
This script outlines the steps take to profile the data from the Food Hygiene Ratings dataset. This script will not make any edits to the data but will note down where cleaning actions should be taken so that this can be performed with an R script in the overall pipeline.

```{r setup, include=FALSE}
# Knitr chunk options
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)

# Knit options
knitr::opts_knit$set(
  root.dir = "../"
)
```

# Packages

Load the packages to be used for this project.
```{r load-packages, echo=TRUE}
library(here)
library(knitr)
library(sf)
library(mapview)
```

# Load the data

The data for profiling is contained in the `fhr_data` object created in the [`2-prep-data.R`](script/2-prep-data.R) script.

```{r load-data}
# Load data
source(here("script/2-prep-data.R"))

# Check that data has loaded
head(fhr_data) %>% 
  kable()
```

# Dataframe profile
The dataframe has `r nrow(fhr_data)` rows and `r ncol(fhr_data)` columns. A `glimpse` of the data is shown below.

```{r glimpse-data}
# Glimpse data
glimpse(fhr_data)
```

# Column profiling
Look at each column in turn and assess the number of unique values. Is any recoding required?

## `fhrsid`
- There are `r length(unique(fhr_data$fhrsid))` unique values.
- This is an ID column with no duplicates.

## `local_authority_business_id`
- There are `r length(unique(fhr_data$local_authority_business_id))` unique values
- This is an ID column with no duplicates.

## `business_name`
- There are `r length(unique(fhr_data$business_name))` unique values.
- Be aware there are some duplicated business names.

## `business_type`
- There are `r length(unique(fhr_data$business_type))` unique values.
```{r business-type-profile}
# What are the unique values?
values_business_type <- fhr_data %>% 
  count(business_type) %>% 
  arrange(desc(n)) 

# Table of values
values_business_type %>% 
  ungroup() %>% 
  mutate(prop = round(n/sum(n),2)) %>% 
  kable()
```

## `business_type_id`
There are `r length(unique(fhr_data$business_type_id))` unique values.
```{r business-type-id-profile}
# What are the unique values?
values_business_type_id <- fhr_data %>% 
  count(business_type_id) %>% 
  arrange(desc(n)) 

# Table of values
values_business_type_id %>% 
  ungroup() %>% 
  mutate(prop = round(n/sum(n),2)) %>% 
  kable()
```

## `rating_value`
- There are `r length(unique(fhr_data$rating_value))` unique values.
```{r rating-value-profile}
# What are the unique values?
values_rating_value <- fhr_data %>% 
  count(rating_value) %>% 
  arrange(desc(n)) 

# Table of values
values_rating_value %>% 
  ungroup() %>% 
  mutate(prop = round(n/sum(n),2)) %>% 
  kable()
```
- Note the mix of character and numeric. Could we remove character?
- Let's also plot a histogram of the numeric values to see the distribution.
```{r rating-value-hist}
# Histogram of rating values
fhr_data %>% 
  select(rating_value) %>% 
  filter(!(rating_value %in% c("Exempt", "AwaitingInspection"))) %>%
  type_convert() %>% 
  pull(rating_value) %>% 
  hist(main = "Distribution of rating values",
       col = "#FF4F5C")
```

## `rating_key`
- There are `r length(unique(fhr_data$rating_key))` unique values.
```{r rating-key-profile}
# What are the unique values?
values_rating_key <- fhr_data %>% 
  count(rating_key) %>% 
  arrange(desc(n)) 

# Table of values
values_rating_key %>% 
  ungroup() %>% 
  mutate(prop = round(n/sum(n),2)) %>% 
  kable()
```
- This column supplements the `rating_value` column nicely. We probably don't need to add another column.

## `rating_date`
- There are `r length(unique(fhr_data$rating_date))` unique values.
- Earliest date is `r min(fhr_data$rating_date, na.rm = TRUE)`
- Latest date is `r max(fhr_data$rating_date, na.rm = TRUE)`
```{r rating-date-profile}
# Summary of values
summary(fhr_data$rating_date)
```

## `local_authority_code`
- There is `r length(unique(fhr_data$local_authority_code))` unique value.
- This should be treated as metadata since the same value applies to every record.

## `local_authority_name`
- There is `r length(unique(fhr_data$local_authority_name))` unique value.
- This should be treated as metadata since the same value applies to every record.

## `local_authority_web_site`
- There is `r length(unique(fhr_data$local_authority_web_site))` unique value.
- This should be treated as metadata since the same value applies to every record.

## `local_authority_email_address`
- There is `r length(unique(fhr_data$local_authority_email_address))` unique value.
- This should be treated as metadata since the same value applies to every record.

## `hygiene`
- There are `r length(unique(fhr_data$hygiene))` unique values.
```{r hygiene-profile}
# What are the unique values?
values_hygiene <- fhr_data %>% 
  count(hygiene) %>% 
  arrange(desc(n)) 

# Table of values
values_hygiene %>% 
  ungroup() %>% 
  mutate(prop = round(n/sum(n),2)) %>% 
  kable()
```
- This column is part of the `score` tag in the original xml file.

## `structural`
- There are `r length(unique(fhr_data$structural))` unique values.
```{r structural-profile}
# What are the unique values?
values_structural <- fhr_data %>% 
  count(structural) %>% 
  arrange(desc(n)) 

# Table of values
values_structural %>% 
  ungroup() %>% 
  mutate(prop = round(n/sum(n),2)) %>% 
  kable()
```
- This column is part of the `score` tag in the original xml file.

## `confidence_in_management`
- There are `r length(unique(fhr_data$confidence_in_management))` unique values.
```{r management-profile}
# What are the unique values?
values_management <- fhr_data %>% 
  count(confidence_in_management) %>% 
  arrange(desc(n)) 

# Table of values
values_management %>% 
  ungroup() %>% 
  mutate(prop = round(n/sum(n),2)) %>% 
  kable()
```
- This column is part of the `score` tag in the original xml file.

## `scheme_type`
- There is `r length(unique(fhr_data$scheme_type))` unique value.
- This should be treated as metadata since the same value applies to every record.

## `new_rating_pending`
- There are `r length(unique(fhr_data$new_rating_pending))` unique values.
```{r pending-profile}
# What are the unique values?
values_pending <- fhr_data %>% 
  count(new_rating_pending) %>% 
  arrange(desc(n)) 

# Table of values
values_pending %>% 
  ungroup() %>% 
  mutate(prop = round(n/sum(n),2)) %>% 
  kable()
```

## `longitude` and `latitude`
- There are `r length(unique(fhr_data$longitude))` unique values of longitude.
- There are `r length(unique(fhr_data$latitude))` unique values of latitude.
- Summary of `longitude`
```{r longitude-profile}
# Summary of values
summary(fhr_data$longitude)
```
- Summary of `latitude`
```{r latitude-profile}
# Summary of values
summary(fhr_data$latitude)
```
- Let's plot the points on a map to ensure they are all valid locations within Waverley.
```{r plot-lat-long}
# Plot the coordinates on a map
fhr_data %>% 
  filter(!is.na(longitude)) %>% 
  st_as_sf(coords = c("longitude", "latitude"),
           crs = 4326) %>% 
  mapview()
```

- Most of the points appear to be in Waverley Borough, which covers the towns of Farnham, Godalming, Haslemere and Cranleigh. However, there are 2 points that appear to outside of Waverley. Clicking on the points reveals more data and we can discover that they refer to `fhrsid` **221504** and **182877**. Let's check out the details of these two records.
```{r review-location-outliers}
# View the address details for the 2 points that appear to be outside of Waverley
fhr_data %>% 
  select(fhrsid,
         business_name,
         business_type,
         longitude,
         latitude,
         address_line1,
         address_line2,
         address_line3,
         address_line4,
         post_code) %>% 
  filter(fhrsid %in% c(221504, 182877)) %>% 
  kable()
```
  - **187877**: The postcode has been entered incorrectly, and thus the lat/long data is incorrect. We can fix this for the analysis.
  - **221504**: The postcode here is correct, but the lat/long coordinates do not match the postcode. We can fix this for the analysis.
  
## `post_code`
- There are `r length(unique(fhr_data$post_codes))` unique values.
- All postcodes in Waverley should start with the characters "GU". Let's check if this is true.
```{r postcode-check}
# Check that all postcodes start with "GU"
fhr_data %>% 
  select(post_code) %>% 
  distinct() %>% 
  filter(!(grepl(pattern = "^GU",
                 x = post_code)))
```
- There are some records where `post_code` is null. This is ok. We'll be excluding those records anyway.
- There are some records where postcode does not start with "GU". Let's see more details about those records.
```{r postcode-outliers}
# Which addresses have non GU postcodes?
fhr_data %>% 
  filter(grepl("^RH", post_code)) %>% 
  select(fhrsid,
         business_name,
         longitude,
         latitude,
         address_line1,
         address_line2,
         address_line3,
         address_line4,
         post_code) %>% 
  kable()
```
- According to the front end [FHR website](https://ratings.food.gov.uk) both of these establishments are covered by Waverley so no need to make any changes.
  
## `fhrs_establishment_id`
- There is `r length(unique(fhr_data$fhrs_establishment_id))` unique value.
- This should be treated as metadata since the same value applies to every record.

# Cleaning requirements identified
- Records with missing location data should be removed. These records refer to establishments that are run from a private address. There are `r fhr_data %>% filter(is.na(longitude)) %>% nrow()` such records in the dataset, representing `r round(((fhr_data %>% filter(is.na(longitude)) %>% nrow()) / nrow(fhr_data))*100)`% of the total.
- `rating_value` has a mix of character and numeric values. Look at keeping a single type in the column (perhaps split data into 2 columns, one for numeric rating and one for rating code.)
- Add the following fields to metadata since the same value is used for every record: `local_authority_code`, `local_authority_name`, `local_authority_web_site`, `local_authority_email_address`, `scheme_type`.
- Append `_score` to the following columns: `hygiene`, `structural`, `confidence_in_management` (abbreviated to `management`).
- `fhrsid` 187877 - the correct postcode for this record should be GU27 2BP. Lat/Long should be 51.0923509197681, -0.7060200868816853 (source: Google Maps using address details).
 - `fhrsid` 221504 - correct lat/long values are 51.21564332626035, -0.8021015970813663 (source: Google Maps using postcode).
